### 2015-03-22 doc/thesis/system.tex
Minor fixes in the text
Adding some missing functions to the section about typing rules

\begin{figure}[!ht]
\textbf{Abstract Syntax}\\
\dstart
$$
\begin{array}{rlr}
s ::= & & \textsc{statements:}\\
& \;\; \mathbf{skip} \;
| \; s_{1} \; ; \; s_{2} & \textit{skip and sequence}\\
& | \; \overline{l} = el & \textit{multiple assignment}\\
& | \; \mathbf{while} \; e \; \mathbf{do} \; s \;
| \; \mathbf{if} \; e \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2} & \textit{control flow}\\
& | \; \mathbf{local} \; \overline{id{:}T} = el \; \mathbf{in} \; s \;
| \; \mathbf{local} \; \overline{id} = el \; \mathbf{in} \; s & \textit{variable declaration}\\
& | \; \mathbf{rec} \; id{:}T = f \; \mathbf{in} \; s & \textit{recursive function} \\
& | \; \mathbf{return} \; el & \textit{return} \\
& | \; \lfloor a \rfloor_{0} & \textit{application with no results}\\
& | \; \mathbf{fun} \; id_{1}{:}id_{2} \; (pl){:}S \; s \;;\; \mathbf{return} \; el & \textit{method declaration}\\
e ::= & & \textsc{expressions:}\\
& \;\; \mathbf{nil} \;
| \; k & \textit{nil and other literals}\\
& | \; id & \textit{variable access}\\
& | \; e_{1}[e_{2}] & \textit{table access}\\
& | \; {<}T{>} \; id & \textit{type coercion}\\
& | \; f & \textit{function declaration}\\
& | \; \{ \; \overline{[e_{1}] = e_{2}} \; \} \;
| \; \{ \; \overline{[e_{1}] = e_{2}},me \; \} & \textit{table constructor}\\
& | \; e_{1} + e_{2} \;
| \; e_{1} \; {..} \; e_{2} \;
| \; e_{1} == e_{2} \;
| \; e_{1} < e_{2} & \textit{binary operations}\\
& | \; e_{1} \;\&\; e_{2} \;
| \; e_{1} \; \mathbf{and} \; e_{2} \;
| \; e_{1} \; \mathbf{or} \; e_{2} & \textit{binary operations}\\
& | \; \mathbf{not} \; e \;
| \; \# \; e & \textit{unary operations} \\
& | \; \lfloor me \rfloor_{1} & \textit{expressions with one result}\\
l ::= & & \textsc{left-hand values:}\\
& \;\; id_{l} & \textit{variable assignment}\\
& | \; e_{1}[e_{2}] & \textit{table assignment}\\
& | \; id[k] \; {<}T{>} & \textit{type coercion}\\
k ::= & & \textsc{literal constants:}\\
& \;\; \mathbf{false} \; | \;
\mathbf{true} \; | \;
{\it int} \; | \;
{\it float} \; | \;
{\it string} & \\
el ::= & & \textsc{expression lists:}\\
& \;\; \overline{e} \; | \;
\overline{e}, me & \\
me ::= & & \textsc{multiple results:}\\
& \;\; a & \textit{application}\\
& | \; {...} & \textit{vararg expression}\\
a ::= & & \textsc{applications:}\\
& \;\; e(el) & \textit{function application}\\
& | \; e{:}n(el) & \textit{method application}\\
f ::= & & \textsc{function declarations:}\\
& \;\; \mathbf{fun} \; (pl){:}S \; s \;;\; \mathbf{return} \; el & \\
pl ::= & & \textsc{parameter lists:}\\
& \;\; \overline{id{:}T} \; | \;
\overline{id{:}T},{...}{:}T & \\
\end{array}
$$
\dend
\caption{Typed Lua abstract syntax}
\label{fig:syntax}
\end{figure}

@@

The function \emph{read} uses the auxiliary function \emph{proj}
to project a union of first-level types, based on an union of
second-level types and an index from a projection type.
In Section \ref{sec:fap} we will discuss how our type system uses
projection types to handle overloaded return types.

@@

This rule uses the auxiliary function \emph{infer} to get the
first-level types of each variable that should be introduced in
the type environment for type checking $s$.
After type checking the statement $s$, rule \textsc{T-LOCAL2} produces a
new type environment $\env_{3}$ without the variables that it introduced
before type checking $s$.

>>>>

\begin{figure}[!ht]
\textbf{Abstract Syntax}\\
\dstart
$$
\begin{array}{rlr}
s ::= & & \textsc{statements:}\\
& \;\; \mathbf{skip} \;
| \; s_{1} \; ; \; s_{2} & \textit{skip and sequence}\\
& | \; \overline{l} = el & \textit{multiple assignment}\\
& | \; \mathbf{while} \; e \; \mathbf{do} \; s \;
| \; \mathbf{if} \; e \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2} & \textit{control flow}\\
& | \; \mathbf{local} \; \overline{id{:}T} = el \; \mathbf{in} \; s \;
| \; \mathbf{local} \; \overline{id} = el \; \mathbf{in} \; s & \textit{variable declaration}\\
& | \; \mathbf{rec} \; id{:}T = f \; \mathbf{in} \; s & \textit{recursive function} \\
& | \; \mathbf{return} \; el & \textit{return} \\
& | \; \lfloor a \rfloor_{0} & \textit{application with no results}\\
& | \; \mathbf{fun} \; id_{1}{:}id_{2} \; (pl){:}S \; s \;;\; \mathbf{return} \; el & \textit{method declaration}\\
e ::= & & \textsc{expressions:}\\
& \;\; \mathbf{nil} \;
| \; k & \textit{nil and other literals}\\
& | \; id & \textit{variable access}\\
& | \; e_{1}[e_{2}] & \textit{table access}\\
& | \; {<}T{>} \; id & \textit{type coercion}\\
& | \; f & \textit{function declaration}\\
& | \; \{ \; \overline{[e_{1}] = e_{2}} \; \} \;
| \; \{ \; \overline{[e_{1}] = e_{2}},me \; \} & \textit{table constructor}\\
& | \; e_{1} + e_{2} \;
| \; e_{1} \; {..} \; e_{2} \;
| \; e_{1} == e_{2} \;
| \; e_{1} < e_{2} & \textit{binary operations}\\
& | \; e_{1} \;\&\; e_{2} \;
| \; e_{1} \; \mathbf{and} \; e_{2} \;
| \; e_{1} \; \mathbf{or} \; e_{2} & \textit{binary operations}\\
& | \; \mathbf{not} \; e \;
| \; \# \; e & \textit{unary operations} \\
& | \; \lfloor me \rfloor_{1} & \textit{expressions with one result}\\
l ::= & & \textsc{left-hand values:}\\
& \;\; id_{l} & \textit{variable assignment}\\
& | \; e_{1}[e_{2}] & \textit{table assignment}\\
& | \; id[k] \; {<}T{>} & \textit{type coercion}\\
k ::= & & \textsc{literal constants:}\\
& \;\; \mathbf{false} \; | \;
\mathbf{true} \; | \;
{\it int} \; | \;
{\it float} \; | \;
{\it string} & \\
el ::= & & \textsc{expression lists:}\\
& \;\; \overline{e} \; | \;
\overline{e}, me & \\
me ::= & & \textsc{multiple results:}\\
& \;\; a & \textit{application}\\
& | \; {...} & \textit{vararg expression}\\
a ::= & & \textsc{applications:}\\
& \;\; e(el) & \textit{function application}\\
& | \; e{:}n(el) & \textit{method application}\\
f ::= & & \textsc{function declarations:}\\
& \;\; \mathbf{fun} \; (pl){:}S \; s \;;\; \mathbf{return} \; el & \\
pl ::= & & \textsc{parameter lists:}\\
& \;\; \overline{id{:}T} \; | \;
\overline{id{:}T},{...}{:}T & \\
\end{array}
$$
\dend
\caption{The abstract syntax of Typed Lua}
\label{fig:syntax}
\end{figure}

@@

The function \emph{read} uses the auxiliary function \emph{proj}
to project a union of first-level types, based on an union of
second-level types and an index from a projection type.
In Section \ref{sec:fap} we will discuss how our type system uses
projection types to handle overloaded return types.
We can define \emph{proj} as follows:
\begin{align*}
proj(T_{1} \times ... \times T_{n}{*}, i) & = \left\{
\begin{array}{ll}
T_{i} & \text{if $i < n$}\\
T_{i} \cup \Nil & \text{if $i = n$}
\end{array} \right.
\\
proj(S_{1} \sqcup S_{2}, i) & = proj(S_{1}, i) \cup proj(S_{2}, i)
\end{align*}

@@

This rule uses the auxiliary function \emph{infer} to get the
most general first-level types of each variable that should be introduced in
the type environment for type checking $s$.
After type checking the statement $s$, rule \textsc{T-LOCAL2} produces a
new type environment $\env_{3}$ without the variables that it introduced
before type checking $s$.
We can define \emph{infer} as follows:
\begin{align*}
infer(T_{1} \times ... \times T_{n}{*}, i) & = \left\{
\begin{array}{ll}
general(T_{i}) & \text{if $i < n$}\\
general(T_{n} \cup \Nil) & \text{if $i >= n$}
\end{array} \right.
\\ \\
general(\False) & = \Boolean\\
general(\True) & = \Boolean\\
general({\it int}) & = \Integer\\
general({\it float}) & = \Number\\
general({\it string}) & = \String\\
general(T_{1} \cup T_{2}) & = general(T_{1}) \cup general(T_{2})\\
general(S_{1} \rightarrow S_{2}) & = general2(S_{1}) \rightarrow general2(S_{2})\\
general(\{K_{1}{:}V_{1}, ..., K_{n}{:}V_{n}\}_{tag}) & = \{K_{1}{:}general(V_{1}), ..., K_{n}{:}general(V_{n})\}_{tag}\\
general(\mu x.T) & = \mu x.general(T)\\
general(T) & = T\\
\\
general2(\Void) & = \Void\\
general2(T{*}) & = general(T){*}\\
general2(T \times P) & = general(T) \times general2(P)\\
general2(S_{1} \sqcup S_{2}) & = general2(S_{1}) \sqcup general2(S_{2})
\end{align*}
